---
title: "part 1 introduction"
author: "sze pui"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.width = 8,

  fig.height = 6,
  out.width = "90%"
)

library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
```

#load data set
```{r}
smoke_df_2020 =read_csv("data/adult20.csv")
smoke_df_2019 =read_csv("data/adult19.csv")
smoke_df_2018 =read_csv("data/adult18.csv")
smoke_df_2017 =read_csv("data/adult17.csv")
smoke_df_2016 =read_csv("data/adult16.csv")
clean_df = smoke_df_2020 %>% 
  #clean variable names
  janitor::clean_names()
```


# Part A The smoking trend in the US over 4 years 
let's clean the dataset for observing the smoking trends in these five years 
lets select the relavant questions first and the following explain the details of the variables
smoking_status: current smoking stastus.


#cleaning data 
```{r}
cleaning_for_smoke_trend= function(dataset){
  dataset%>% 
  janitor::clean_names()%>% 
  select(cignow_a, smkcigst_a,sex_a,hispallp_a, agep_a, smkev_a)%>% 
  rename(smoke_freq = cignow_a ,over100= smkev_a,  smoking_status = smkcigst_a, sex = sex_a,race =hispallp_a , age =  agep_a)}
#select(cignow_a, smkcigst_a,smkage_a)
#reminder for later: smoking_age =smkage_a 
smoke_trend_2020 = cleaning_for_smoke_trend(smoke_df_2020)
smoke_trend_2019  = cleaning_for_smoke_trend(smoke_df_2019)

#cleaning data for 2018, 2017,2016 
#rewrite the function since they have a different variable code
cleaning_for_smoke_trend_2 = function(dataset){
  dataset%>% 
  janitor::clean_names()%>% 
  select(smkstat2,sex, age_p, racerpi2,smkev,smkreg)%>% 
  rename(smoking_status =smkstat2,over100= smkev, age = age_p, race= racerpi2, smokeage =smkreg)}

smoke_trend_2018 = cleaning_for_smoke_trend_2(smoke_df_2018)
smoke_trend_2017 = cleaning_for_smoke_trend_2(smoke_df_2017)
```

#Plot of number of people ever smoked over 100 cigarettes over the past 4 years 
```{r}
over_100 = function(df){
  df%>% 
  mutate(over100 = as.factor(over100),
         over100 = recode(over100,"1" = "yes", "2" ="no"))%>% 
  filter(over100 =="yes"|over100 =="no") %>% 
  count(over100)%>% 
  mutate(ans_sum = nrow(df), 
          over100percentage = n/ans_sum)}
  

a= over_100(smoke_trend_2020)%>%
  mutate(year =  "2020") 
b= over_100(smoke_trend_2019 )%>%
  mutate(year = "2019") 
c = over_100(smoke_trend_2018)%>%
  mutate(year =  "2018")
d = over_100(smoke_trend_2017)%>%
  mutate(year =  "2017") 

plot_df0= bind_rows(a,b,c,d)%>%
  select(-n, -ans_sum)%>%
   ggplot(aes(fill = over100, y=over100percentage,x = year)) + 
    geom_bar(position="dodge", stat="identity")+labs(
    title = "Have you ever smoked at least 100 cigarettes in your ENTIRE LIFE?",
    x = "year",
    y = "percentage of people",
    caption = "Data from National Health Survey")
```

#Plot of the changing trend of people with different smoking status
```{r}
 finding_smoking_trend = function(dataframe){
   dataframe %>% 
  select(smoking_status)%>%
  mutate(smoking_status = recode(smoking_status,"1" = "current every day smoker", "2" ="current some day smoker",  "3"= "former smoker","4" = "never smoker"))%>% 
  count(smoking_status)%>%
     mutate(ans_sum = nrow(dataframe),
            percentage = n/ans_sum)%>%
  drop_na()}

#apply the function to know the summary of smoking status 
#can i use a map function here?
#have to add year and undergo pivot_longer
#problem encountered: temporarily there is no data available in 2016
smoke_trend_20 = smoke_trend_2020 %>% 
  finding_smoking_trend()%>%
  mutate(year =  "2020") 
        
smoke_trend_19 = smoke_trend_2019 %>% 
  finding_smoking_trend()%>%
  mutate(year =  "2019")

 smoke_trend_18= smoke_trend_2018 %>% 
  finding_smoking_trend()%>%
  mutate(year =  "2018")

smoke_trend_17 = smoke_trend_2017 %>% 
  finding_smoking_trend()%>%
  mutate(year =  "2017")

smoke_trend_overall = bind_rows(smoke_trend_17,smoke_trend_18,smoke_trend_19,smoke_trend_20)%>%
  rename(ppl_sum = n)%>%
  relocate(year)

#lets create a ggpplot
# ***task： try it to make it interactive soon....
#Problem: Should I use rate instead....???
smoke_trend_overall %>% 
  ggplot(aes(x = year, y = percentage, group = smoking_status)) + 
  geom_point(aes(color = smoking_status))+ geom_line(aes(color = smoking_status))+ 
 ggtitle("Overall smoking trend of the U.S in recent 4 years according to NHIS ")+
  labs(y = "percentage", caption = "Data from National Health Survey")

#ggplotly(smoke_trend_overall_plot)

```



