---
title: "Overall Smoking Trend"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: sandstone
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
library(viridis)
library(gganimate)
library(gifski)
```

In this study, we filtered the smokers from all respondents of the National Health Interview Survey and plotted graphs to observe the distribution of smokers by sex, age, and race groups.

```{r, include=FALSE}
#load data set
smoke_df_2020 =read_csv("data/adult20.csv")
smoke_df_2019 =read_csv("data/adult19.csv")
smoke_df_2018 =read_csv("data/adult18.csv")
smoke_df_2017 =read_csv("data/adult17.csv")
smoke_df_2016 =read_csv("data/adult16.csv")
clean_df = smoke_df_2020 %>% 
  #clean variable names
  janitor::clean_names()
```



```{r, include=FALSE}
# Part A The smoking trend in the US over 4 years 
#let's clean the dataset for observing the smoking trends in these five years 
#lets select the relavant questions first and the following explain the details of the variables
#smoking_status: current smoking stastus.
#cleaning data 

cleaning_for_smoke_trend= function(dataset){
  dataset%>% 
  janitor::clean_names()%>% 
  select(cignow_a, smkcigst_a,sex_a,hispallp_a, agep_a, smkev_a)%>% 
  rename(smoke_freq = cignow_a ,over100= smkev_a,  smoking_status = smkcigst_a, sex = sex_a,race =hispallp_a , age =  agep_a)}
#select(cignow_a, smkcigst_a,smkage_a)
#reminder for later: smoking_age =smkage_a 
smoke_trend_2020 = cleaning_for_smoke_trend(smoke_df_2020)
smoke_trend_2019  = cleaning_for_smoke_trend(smoke_df_2019)

#cleaning data for 2018, 2017,2016 
#rewrite the function since they have a different variable code
cleaning_for_smoke_trend_2 = function(dataset){
  dataset%>% 
  janitor::clean_names()%>% 
  select(smkstat2,sex, age_p, racerpi2,smkev,smkreg)%>% 
  rename(smoking_status =smkstat2,over100= smkev, age = age_p, race= racerpi2, smokeage =smkreg)}

smoke_trend_2018 = cleaning_for_smoke_trend_2(smoke_df_2018)
smoke_trend_2017 = cleaning_for_smoke_trend_2(smoke_df_2017)
```

```{r,include= FALSE}
## Plot of number of people ever smoked over 100 cigarettes over the past 4 years 

over_100 = function(df){
  df%>% 
  mutate(over100 = as.factor(over100),
         over100 = recode(over100,"1" = "yes", "2" ="no"))%>% 
  filter(over100 =="yes"|over100 =="no") %>% 
  count(over100)%>% 
  mutate(ans_sum = nrow(df), 
          over100percentage = n/ans_sum)}
  

a= over_100(smoke_trend_2020)%>%
  mutate(year =  "2020") 
b= over_100(smoke_trend_2019 )%>%
  mutate(year = "2019") 
c = over_100(smoke_trend_2018)%>%
  mutate(year =  "2018")
d = over_100(smoke_trend_2017)%>%
  mutate(year =  "2017") 

plot0 = bind_rows(a,b,c,d)%>%
  select(-n, -ans_sum)%>%
  mutate(over100percentage = over100percentage*100) %>% 
   ggplot(aes(fill = over100, y=over100percentage,x = year)) + 
    geom_bar(position="dodge", stat="identity")+labs(
    title = "Have you ever smoked at least 100 cigarettes in your ENTIRE LIFE?",
    x = "year",
    y = "percentage of people",
    caption = "Data from National Health Survey")

ggplotly(plot0 )


#For year between 2017 and 2020, the percentage of people answering "no" was much higher than that of people #answering "yes". There is no obvious percentage change in the proportion of people smoking more than 100 cigarettes #vs people who are not in the past 4 years. The ratio of percentage in people smoking more than 100 versus to people #who are not is always 6:4.
```


### Overall smoking trend from 2017 to 2020

```{r, include = FALSE}
 finding_smoking_trend = function(dataframe){
   dataframe %>% 
  select(smoking_status)%>%
  mutate(smoking_status = recode(smoking_status,"1" = "current every day smoker", "2" ="current some day smoker",  "3"= "former smoker","4" = "never smoker"))%>% 
  count(smoking_status)%>%
     mutate(ans_sum = nrow(dataframe),
            percentage = n/ans_sum)%>%
  drop_na()}

#apply the function to know the summary of smoking status 
#can i use a map function here?
#have to add year and undergo pivot_longer
#problem encountered: temporarily there is no data available in 2016
smoke_trend_20 = smoke_trend_2020 %>% 
  finding_smoking_trend()%>%
  mutate(year =  "2020") 
        
smoke_trend_19 = smoke_trend_2019 %>% 
  finding_smoking_trend()%>%
  mutate(year =  "2019")

 smoke_trend_18= smoke_trend_2018 %>% 
  finding_smoking_trend()%>%
  mutate(year =  "2018")

smoke_trend_17 = smoke_trend_2017 %>% 
  finding_smoking_trend()%>%
  mutate(year =  "2017")

smoke_trend_overall = bind_rows(smoke_trend_17,smoke_trend_18,smoke_trend_19,smoke_trend_20)%>%
  rename(ppl_sum = n)%>%
  relocate(year)

```

```{r, warning = FALSE, message = FALSE}
#lets create a ggpplot
# ***task： try it to make it interactive soon....
#Problem: Should I use rate instead....???
smoke_trend_overall_plot = smoke_trend_overall %>% 
  mutate(year = as.numeric(year),
         percentage = percentage*100) %>% 
  ggplot(aes(x = year, y = percentage, group = smoking_status)) + 
  geom_point(aes(color = smoking_status))+ geom_line(aes(color = smoking_status))+ 
 ggtitle("Overall Smoking Trend of the U.S in Recent 4 Years According to NHIS ")+
  labs(y = "Percentage", caption = "Data from National Health Survey")+
  transition_reveal(year)

animate(smoke_trend_overall_plot , renderer = gifski_renderer())

```

There is a  slightly increase in the percentage of former smoker and never smoker from 2017 to 2020. The percentage of former smoker increased from 24.6% to 25.8%. The percentage of never smoker increased from 59.9% to 60.9%. Meanwhile, the percentage of current every day smoker and current some day smoker were slightly decreased by 2.0% and 0.7% respectively.

```{r,include=FALSE}
#Group by race, age and sex to see their proportion among the smokers

#cleaning data 
clean_df = function(df){
  df %>%
  filter(smoking_status == "1" |smoking_status == "2") %>% 
  mutate(sex = recode(sex,"1" = "male", "2" ="female"), 
         race =recode(race, "1" = "hispanic ", "2" ="white" , "3"="black/african american", "4"= "asian ", "5" = "aian", "6" = "aian and any other group", "7" ="other races"),
         smoking_status =  case_when(smoking_status == "1"|smoking_status == "2"~ "smoker"), 
          agegroup =  case_when(
    between(age, 18, 25) ~ "18-25",
    between(age, 26, 35) ~ "26-35",
    between(age, 36, 45) ~ "36-45",
    between(age, 46, 55) ~ "46-55",
    between(age, 56, 65) ~ "56-65",
    between(age, 66, 100) ~ "65+",
    TRUE ~ NA_character_),
   sex = as.factor(sex),
   race = as.factor(race),
   smoker_sum = nrow(df)) %>% 
  select(-age)} %>% 
  group_by(sex) %>% 
  mutate(sex_count = n())

```

### Overall distribution of smokers among different sex groups from 2017 to 2020

```{r,include=FALSE}
a1 =clean_df (smoke_trend_2020) %>% 
  mutate(year= "2020")

b1 = clean_df (smoke_trend_2019) %>% 
  mutate(year= "2019")

c1 = clean_df (smoke_trend_2018) %>% 
  mutate(year= "2018")

d1 = clean_df (smoke_trend_2017) %>% 
  mutate(year= "2017")
```

```{r, warning = FALSE, message = FALSE}
#smokers' sex proportion
plot1 = bind_rows(a1,b1,c1,d1) %>% 
  count(year, sex) %>% 
  drop_na() %>% 
  ggplot(aes(fill=sex, y=n, x=year))  + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_viridis(discrete = T)+
    labs(title = "The Distribution of smokers by sex over the 4 years",y = "Smoker sum", caption = "Data from National Health Survey") 

ggplotly(plot1)
```
<<<<<<< HEAD
=======

**!!Interesting fact** : There is more male smokers than the female smokers!
In the past 4 years, it can be seen that the total sum of male smokers is higher that of female smoker. 
>>>>>>> c20c41cf73075c7aed9a659cc208fb4b72768845

The plot shows that there were more male smokers than female smokers every year of the past four years.

### Overall distribution of smokers among different age groups from 2017 to 2020

```{r, warning = FALSE, message = FALSE}
#smoker's age proportion 
bind_rows(a1,b1,c1,d1) %>% 
  count(year, agegroup) %>% 
  drop_na() %>% 
  ggplot(aes(fill=agegroup, y=n, x=year))  + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_viridis(discrete = T)+
    labs(title = "The Distribution of smokers by age group over the 4 years",y = "Smoker sum", caption = "Data from National Health Survey") 
```

The age group 56-65 has the most smokers every year in the past four years. The age group 36-45 and 46-55 are slightly lower than 56-55 but worth noting as well. The age group 18-25 has the least smokers. There was little difference in the age distribution of smokers from 2017 to 2020. Smokers in the 18-25 age group showed a decrease after 2017. This is a good sign.


```{r,include = FALSE}
#smoker's race proportion
plot3 = bind_rows(a1,b1,c1,d1) %>% 
  count(year, race) %>% 
  drop_na() %>% 
  ggplot(aes(fill=race, y=n, x=year))  + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_viridis(discrete = T)+
    labs(title = "Total smoker sum in different race groups over the  past 4 years",y = "smoker sum", caption = "Data from National Health Survey") 


#This is a graph showing total sum of smokers among different race.
#This can only be a reference since the population with respective to each race is not taken into consideration.
#In the following graph the proportion of smokers among different race, sex and agegroup will be shown

ggplotly(plot3)
```


### Overall comparison of smoking proportion among people in different age, sex, and race groups from 2017 to 2020

We compared the smoker proportion of each year among different age, sex, and race groups. The smoker proportion is obtained by diving the number of smokers to the total sum of people in the respective group.

```{r, include = FALSE}
  finding_smoking_dis =  function(dataframe){
   dataframe %>% 
 select(smoking_status, race, age, sex)%>%
  filter(sex== "1" | sex== "2")%>%
  subset(race != "97"|race != "98"|race != "99") %>% 
  mutate(sex = as.factor(sex),
         race = as.factor(race)) %>% 
  mutate(sex = recode(sex,"1" = "male", "2" ="female"), 
         race =recode(race, "1" = "hispanic ", "2" ="white" , "3"="black/african american", "4"= "asian ", "5" = "aian", "6" = "aian and any other group", "7" ="other races"),
         smoking_status = 
           case_when(smoking_status == "1"|smoking_status == "2"~ "smoker",
                     smoking_status == "3"|smoking_status == "4"~ "non_smoker"), 
          agegroup =  case_when(
    between(age, 18, 25) ~ "18-25",
    between(age, 26, 35) ~ "26-35",
    between(age, 36, 45) ~ "36-45",
    between(age, 46, 55) ~ "46-55",
    between(age, 56, 65) ~ "56-65",
    between(age, 66, 100) ~ "65+",
    TRUE ~ NA_character_),
    agegroup = as.factor(agegroup)) %>%
  select(-age)%>%
  filter(smoking_status == "smoker"| smoking_status == "non_smoker") %>% 
  group_by(smoking_status, sex,agegroup,race)%>%
  count()%>%
  pivot_wider(
    names_from = "smoking_status",
    values_from = n)%>%
  drop_na()%>%
  mutate(percentage = smoker/(smoker + non_smoker)) %>% 
  select(-smoker, -non_smoker)%>%  
  ggplot(aes(x=agegroup, y=percentage, size=percentage, color=race)) +
   #there are some repetitive in the color 
    geom_point(alpha=0.5) +
    scale_size(range = c(.1, 13), name="Smoking proportion")+ facet_wrap(.~sex)+  labs(caption = "Data from National Health Survey")}
```


```{r, warning = FALSE, message = FALSE}
#Plot 2020
plot4 = finding_smoking_dis(smoke_trend_2020)+labs(title = " Distribution and Proportion of Smokers grouped by age,race and sex in year 2020", y = "proportion", caption = "Data from National Health Survey") + theme(axis.text.x = element_text(angle = 60, vjust = 0.5, hjust=1))
ggplotly(plot4)
```

In 2020, it can be seen that there is a high proportion of smokers in AIAN/ AIAN and any other group. Meanwhile, Asian group, especially female, has a comparatively low smokers proportion.

```{r, warning = FALSE, message = FALSE}
#Plot 2019 
plot5 = finding_smoking_dis(smoke_trend_2019)+labs(title = " Distribution and Proportion of Smokers grouped by age,race and sex in year 2019", y = "proportion", caption = "Data from National Health Survey") + theme(axis.text.x = element_text(angle = 60, vjust = 0.5, hjust=1))
ggplotly(plot5)
```

In 2019, the smoking proportion of male AIAN and other group is the highest in all age group except 65+.
The smoker proportion in male AIAN and other group aged 56-65 is especially prominent, which reached to 48.3%, meaning that nearly half of the group are smokers. 
Meanwhile, Asian female has the lowest smoking proportion among all age groups.

```{r, warning = FALSE, message = FALSE}
#Plot 2018 
plot6 = finding_smoking_dis(smoke_trend_2018)+labs(title = " Distribution of smoking among agegroup and race in year 2018", y = "proportion", caption = "Data from National Health Survey") + theme(axis.text.x = element_text(angle = 60, vjust = 0.5, hjust=1))
ggplotly(plot6)
```

In 2018, the smoker proportion in female AIAN aged 46-55 is especially prominent, which reached to 60%, meaning that more than half of the group are smokers. There is a slightly increasing trend in smokers proportion in female Black/African American along with the increase of age group between 18-25 to 56-65.

```{r, warning = FALSE, message = FALSE}
#Ploto 2017 
plot7 = finding_smoking_dis(smoke_trend_2017)+labs(title = " Distribution and Proportion of Smokers grouped by age,race and sex in year 2017", y = "proportion", caption = "Data from National Health Survey") + theme(axis.text.x = element_text(angle = 60, vjust = 0.5, hjust=1))

ggplotly(plot7)
```

In 2017, there is a comparatively low smoking proportion in both female and male Asian group among all age group.

Overall, the smoking proportion of Asians, especially females is the least among age group.
