---
title: "part 1 introduction"
author: "sze pui"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
library(viridis)
```

#load data set
```{r}
smoke_df_2020 =read_csv("data/adult20.csv")
smoke_df_2019 =read_csv("data/adult19.csv")
smoke_df_2018 =read_csv("data/adult18.csv")
smoke_df_2017 =read_csv("data/adult17.csv")
smoke_df_2016 =read_csv("data/adult16.csv")
clean_df = smoke_df_2020 %>% 
  #clean variable names
  janitor::clean_names()
```


# Part A The smoking trend in the US over 4 years 
let's clean the dataset for observing the smoking trends in these five years 
lets select the relavant questions first and the following explain the details of the variables
smoking_status: current smoking stastus.


#cleaning data 
```{r}
cleaning_for_smoke_trend= function(dataset){
  dataset%>% 
  janitor::clean_names()%>% 
  select(cignow_a, smkcigst_a,sex_a,hispallp_a, agep_a, smkev_a)%>% 
  rename(smoke_freq = cignow_a ,over100= smkev_a,  smoking_status = smkcigst_a, sex = sex_a,race =hispallp_a , age =  agep_a)}
#select(cignow_a, smkcigst_a,smkage_a)
#reminder for later: smoking_age =smkage_a 
smoke_trend_2020 = cleaning_for_smoke_trend(smoke_df_2020)
smoke_trend_2019  = cleaning_for_smoke_trend(smoke_df_2019)

#cleaning data for 2018, 2017,2016 
#rewrite the function since they have a different variable code
cleaning_for_smoke_trend_2 = function(dataset){
  dataset%>% 
  janitor::clean_names()%>% 
  select(smkstat2,sex, age_p, racerpi2,smkev,smkreg)%>% 
  rename(smoking_status =smkstat2,over100= smkev, age = age_p, race= racerpi2, smokeage =smkreg)}

smoke_trend_2018 = cleaning_for_smoke_trend_2(smoke_df_2018)
smoke_trend_2017 = cleaning_for_smoke_trend_2(smoke_df_2017)
```

#Plot of number of people ever smoked over 100 cigarettes over the past 4 years 
```{r}
over_100 = function(df){
  df%>% 
  mutate(over100 = as.factor(over100),
         over100 = recode(over100,"1" = "yes", "2" ="no"))%>% 
  filter(over100 =="yes"|over100 =="no") %>% 
  count(over100)%>% 
  mutate(ans_sum = nrow(df), 
          over100percentage = n/ans_sum)}
  

a= over_100(smoke_trend_2020)%>%
  mutate(year =  "2020") 
b= over_100(smoke_trend_2019 )%>%
  mutate(year = "2019") 
c = over_100(smoke_trend_2018)%>%
  mutate(year =  "2018")
d = over_100(smoke_trend_2017)%>%
  
  mutate(year =  "2017") 

plot_df0= bind_rows(a,b,c,d)%>%
  select(-n, -ans_sum)%>%
   ggplot(aes(fill = over100, y=over100percentage,x = year)) + 
    geom_bar(position="dodge", stat="identity")+labs(
    title = "Have you ever smoked at least 100 cigarettes in your ENTIRE LIFE?",
    x = "year",
    y = "percentage of people",
    caption = "Data from National Health Survey")
```


#Overall distribution of smokers among different age, sex and race group
```{r}
 finding_smoking_trend = function(dataframe){
   dataframe %>% 
  select(smoking_status)%>%
  mutate(smoking_status = recode(smoking_status,"1" = "current every day smoker", "2" ="current some day smoker",  "3"= "former smoker","4" = "never smoker"))%>% 
  count(smoking_status)%>%
     mutate(ans_sum = nrow(dataframe),
            percentage = n/ans_sum)%>%
  drop_na()}

#apply the function to know the summary of smoking status 
#can i use a map function here?
#have to add year and undergo pivot_longer
#problem encountered: temporarily there is no data available in 2016
smoke_trend_20 = smoke_trend_2020 %>% 
  finding_smoking_trend()%>%
  mutate(year =  "2020") 
        
smoke_trend_19 = smoke_trend_2019 %>% 
  finding_smoking_trend()%>%
  mutate(year =  "2019")

 smoke_trend_18= smoke_trend_2018 %>% 
  finding_smoking_trend()%>%
  mutate(year =  "2018")

smoke_trend_17 = smoke_trend_2017 %>% 
  finding_smoking_trend()%>%
  mutate(year =  "2017")

smoke_trend_overall = bind_rows(smoke_trend_17,smoke_trend_18,smoke_trend_19,smoke_trend_20)%>%
  rename(ppl_sum = n)%>%
  relocate(year)

#lets create a ggpplot
# ***task： try it to make it interactive soon....
#Problem: Should I use rate instead....???
smoke_trend_overall %>% 
  ggplot(aes(x = year, y = percentage, group = smoking_status)) + 
  geom_point(aes(color = smoking_status))+ geom_line(aes(color = smoking_status))+ 
 ggtitle("Overall smoking trend of the U.S in recent 4 years according to NHIS ")+
  labs(y = "percentage", caption = "Data from National Health Survey")

#ggplotly(smoke_trend_overall_plot)

```

#Group by race, age and sex to see their proportion among the smokers
```{r}
#cleaning data 
clean_df = function(df){
  df %>%
  filter(smoking_status == "1" |smoking_status == "2") %>% 
  mutate(sex = recode(sex,"1" = "male", "2" ="female"), 
         race =recode(race, "1" = "hispanic ", "2" ="white" , "3"="black/african american", "4"= "asian ", "5" = "aian", "6" = "aian and any other group", "7" ="other races"),
         smoking_status =  case_when(smoking_status == "1"|smoking_status == "2"~ "smoker"), 
          agegroup =  case_when(
    between(age, 18, 25) ~ "18-25",
    between(age, 26, 35) ~ "26-35",
    between(age, 36, 45) ~ "36-45",
    between(age, 46, 55) ~ "46-55",
    between(age, 56, 65) ~ "56-65",
    between(age, 66, 100) ~ "65+",
    TRUE ~ NA_character_),
   sex = as.factor(sex),
   race = as.factor(race),
   smoker_sum = nrow(df)) %>% 
  select(-age)} %>% 
  group_by(sex) %>% 
  mutate(sex_count = n())
```

```{r}
a1 =clean_df (smoke_trend_2020) %>% 
  mutate(year= "2020")

b1 = clean_df (smoke_trend_2019) %>% 
  mutate(year= "2019")

c1 = clean_df (smoke_trend_2018) %>% 
  mutate(year= "2018")

d1 = clean_df (smoke_trend_2017) %>% 
  mutate(year= "2017")

#smokers' sex proportion
bind_rows(a1,b1,c1,d1) %>% 
  count(year, sex) %>% 
  drop_na() %>% 
  ggplot(aes(fill=sex, y=n, x=year))  + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_viridis(discrete = T)+
    labs(title = "smokers' sex proportion in  past 4 years",y = "smoker sum", caption = "Data from National Health Survey") 

#plotly

#smoker's age proportion 
bind_rows(a1,b1,c1,d1) %>% 
  count(year, agegroup) %>% 
  drop_na() %>% 
  ggplot(aes(fill=agegroup, y=n, x=year))  + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_viridis(discrete = T)+
    labs(title = "total smokers'in different age groups in  past 4 years",y = "smoker sum", caption = "Data from National Health Survey") 

#smoker's race proportion
bind_rows(a1,b1,c1,d1) %>% 
  count(year, race) %>% 
  drop_na() %>% 
  ggplot(aes(fill=race, y=n, x=year))  + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_viridis(discrete = T)+
    labs(title = "total smokers'in different race groups in  past 4 years",y = "smoker sum", caption = "Data from National Health Survey") 
```



```{r}
  finding_smoking_dis =  function(dataframe){
   dataframe %>% 
 select(smoking_status, race, age, sex)%>%
  filter(sex== "1" | sex== "2")%>%
  subset(race != "97"|race != "98"|race != "99") %>% 
  mutate(sex = as.factor(sex),
         race = as.factor(race)) %>% 
  mutate(sex = recode(sex,"1" = "male", "2" ="female"), 
         race =recode(race, "1" = "hispanic ", "2" ="white" , "3"="black/african american", "4"= "asian ", "5" = "aian", "6" = "aian and any other group", "7" ="other races"),
         smoking_status = 
           case_when(smoking_status == "1"|smoking_status == "2"~ "smoker",
                     smoking_status == "3"|smoking_status == "4"~ "non_smoker"), 
          agegroup =  case_when(
    between(age, 18, 25) ~ "18-25",
    between(age, 26, 35) ~ "26-35",
    between(age, 36, 45) ~ "36-45",
    between(age, 46, 55) ~ "46-55",
    between(age, 56, 65) ~ "56-65",
    between(age, 66, 100) ~ "65+",
    TRUE ~ NA_character_),
    agegroup = as.factor(agegroup)) %>%
  select(-age)%>%
  filter(smoking_status == "smoker"| smoking_status == "non_smoker") %>% 
  group_by(smoking_status, sex,agegroup,race)%>%
  count()%>%
  pivot_wider(
    names_from = "smoking_status",
    values_from = n)%>%
  drop_na()%>%
  mutate(percentage = smoker/(smoker + non_smoker)) %>% 
  select(-smoker, -non_smoker)%>%  
  ggplot(aes(x=agegroup, y=percentage, size=percentage, color=race)) +
   #there are some repetitive in the color 
    geom_point(alpha=0.5) +
    scale_size(range = c(.1, 13), name="Smoking proportion(%)")+ facet_wrap(.~sex)+  labs(caption = "Data from National Health Survey")}

#Plot 2020
finding_smoking_dis(smoke_trend_2020)+labs(title = " Distribution of smoking among agegroup and race in year 2020")
#Plot 2019 
finding_smoking_dis(smoke_trend_2019)+labs(title = " Distribution of smoking among agegroup and race in year 2019")
#Plot 2018 
finding_smoking_dis(smoke_trend_2018)+labs(title = " Distribution of smoking among agegroup and race in year 2018")
#Ploto 2017 
finding_smoking_dis(smoke_trend_2017)+labs(title = " Distribution of smoking among agegroup and race in year 2017")

#ggplotly
```



